---
title: "costofemp041523"
author: "Caroline Motzer"
date: "2023-04-15"
output: pdf_document
---
##Setting working directory
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Volumes/BoorWiedmannLab/clm254_CarolineMotzer/Cost of EMP project/Data")
```
##loading packages
```{r}
library(rlang)
library(readr)
library(car)
library(emmeans)
library(ggplot2)
na.rm = TRUE
```

##loading in Data and setting working directory
```{r}
costofemp_redit_final <- read_csv("/Volumes/BoorWiedmannLab/clm254_CarolineMotzer/Cost of EMP project/Data/costofemp_redit_final.csv")
View(costofemp_redit_final)
data <- costofemp_redit_final
```
##Creating model to test association between emp investment cost, estimated total value of finished product, EMP dedication (Fscs) and overall prevalence
```{r}
costofemp2<- lm(log(EMPinvestmentcost)~ esttotalvaluefinished + Fscs + overallprev2 + annualmilksolidsproc, data = data)
summary(costofemp2)
```
##Checking model assumptions and double checking interpretation (via anova)
```{r}
plot(costofemp2)
hist(costofemp2$residuals)
plot(predict(costofemp2), residuals(costofemp2))
plot(data$esttotalvaluefinished, data$EMPinvestmentcost)
plot(data$overallprev2, data$EMPinvestmentcost)
plot(data$Fscs, data$EMPinvestmentcost)
vif(costofemp2)
anova(costofemp2)
```

##checking correlation between variables
```{r}
cor.test(data$annualmilksolidsproc, data$EMPinvestmentcost)
cor.test(data$overallprev2, data$EMPinvestmentcost)
```
##setting up model to see if change in prevalence from start of year to end of year is associated with investment on corrective actions
```{r}
costofemp_redit_final$correctiveratio<- costofemp_redit_final$correctiveactionscost/costofemp_redit_final$EMPinvestmentcost
hist(costofemp_redit_final$correctiveratio, breaks=40)
costofemp_redit_final$correctivemajority<- "no"
  costofemp_redit_final$correctivemajority [costofemp_redit_final$correctiveratio>0.8]<-"yes"
costofemp_redit_final$prevdif<-costofemp_redit_final$Validprev-costofemp_redit_final$Initialprev
correctiveactionmodel3<- lm(costofemp_redit_final$prevdif~costofemp_redit_final$correctiveratio)
summary(correctiveactionmodel3)
View(costofemp_redit_final)
```
##identifying correlation between plant's perception of pathogen control and final listeria prevalence
```{r}
cor.test(costofemp_redit_final$perceived_control_past, costofemp_redit_final$Validprev)
plot(costofemp_redit_final$perceived_control_past, costofemp_redit_final$Validprev)
```
